---
title: "Homework 3"
author: Megan Marziali
output: github_document
---

# Beginning infrastructure

This code chunk calls any relevant libraries.

```{r, message = FALSE}
library(tidyverse)
library(p8105.datasets)
```


## Problem 1

#### Loading instacart dataset

The following code chunk loads the Instacart dataset

```{r}
data("instacart")
```

The variables included in the instacart dataset are `r names(instacart)`. There are a total of `r nrow(instacart)` observations, and `r ncol(instacart)` variables.

#### Answering specific questions

The following code is to check how many aisles.

## Problem 2

The following code is to load in the accelerometer data.

```{r}
accel_df = 
  read_csv("./problem2/accel_data.csv",
           na = "") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "activity",
    names_prefix = "activity_",
    values_to = "measures"
  ) %>% 
  mutate(
    day_type = recode(
      day, 
      Monday = "Weekday",
      Tuesday = "Weekday",
      Wednesday = "Weekday",
      Thursday = "Weekday",
      Friday = "Weekday", 
      Saturday = "Weekend",
      Sunday = "Weekend"),
    day = factor(day),
    day = forcats::fct_relevel(day, c("Monday",
                                      "Tuesday",
                                      "Wednesday",
                                      "Thursday",
                                      "Friday",
                                      "Saturday",
                                      "Sunday"))
  )
```


The following code makes a table.

```{r}
accel_df %>% 
  group_by(day_id) %>% 
  summarize(
    total_activity = sum(measures)
  )
```




