---
title: "Homework 3"
author: Megan Marziali
output: github_document
---

## Beginning infrastructure

This code chunk calls any relevant libraries and setting options.

```{r, message = FALSE}
library(tidyverse)
library(p8105.datasets)
library(ggridges)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = 0.6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Problem 1

#### Loading instacart dataset

The following code chunk loads the Instacart dataset

```{r}
data("instacart")
```

The variables included in the instacart dataset are `r names(instacart)`. There are a total of `r nrow(instacart)` observations, and `r ncol(instacart)` variables.

#### Answering specific questions

The following code is to check how many aisles.

## Problem 2

The following code is to load in the accelerometer data.

```{r message=FALSE}
accel_df = 
  read_csv("./problem2/accel_data.csv",
           na = "") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "activity",
    names_prefix = "activity_",
    values_to = "measures"
  ) %>% 
  mutate(
    day_type = recode(
      day, 
      Monday = "Weekday",
      Tuesday = "Weekday",
      Wednesday = "Weekday",
      Thursday = "Weekday",
      Friday = "Weekday", 
      Saturday = "Weekend",
      Sunday = "Weekend"),
    day = factor(day),
    day = forcats::fct_relevel(day, c("Monday",
                                      "Tuesday",
                                      "Wednesday",
                                      "Thursday",
                                      "Friday",
                                      "Saturday",
                                      "Sunday")),
    activity = as.integer(activity)
  )
```

This dataset includes accelerometer data collected from a 65 year old patient. The variables included are `r names(accel_df)`. There are a total of `r nrow(accel_df)` rows and `r ncol(accel_df)` columns. The total number of **activity** measurements taken is `r max(pull(accel_df, activity))`. Measurements were taken every `r levels(pull(accel_df, day))`, for a total of `r max(pull(accel_df, week))` weeks and `r max(pull(accel_df, day_id))` days.

The value of the accelerometer readings as per the **measures** variable ranges from `r min(pull(accel_df, measures))` to `r max(pull(accel_df, measures))`. The mean of the accelerometer readings is `r mean(pull(accel_df, measures), n = 2)`, with a standard deviation of `r sd(pull(accel_df, measures), n = 2)`. The median is `r median(pull(accel_df, measures), n = 2)`, with an IQR of `r IQR(pull(accel_df, measures), n = 2)`. 

### Aggregating data

The following code makes a table, aggregated by total activity per day.

```{r message=FALSE}
accel_agg = 
  accel_df %>% 
  group_by(day) %>% 
  summarize(
    total_activity = sum(measures)
  ) %>% 
  knitr::kable(digits = 1)
```

It seems that the patient is the least active on Saturdays, and the most active on Fridays. The patient is somewhat active on Mondays, Tuesdays, and Sundays but more active on Wednesdays, Thursdays and Fridays.

```{r message = FALSE}
accel_df %>% 
  group_by(day_id, day) %>% 
  summarize(
    activity_time = sum(measures)
  ) %>% 
  ggplot(aes(x = activity_time, color = day)) +
  geom_density()
```

## Problem 3

The following code loads in the appropriate dataset.

```{r}
data("ny_noaa")
```

The following code cleans and tidies the data.



